---
title: "SPRIS Proj 4 Draft"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(tidyverse)
library(haven)
library(gtsummary)
library(KMsurv)
library(survival)
library(survival)
library(survminer)
library(ggplot2)
library(patchwork)
library(pec)
library(lme4)
library(ggsurvfit)
```

# Problem 1

```{r}
baseline = read_excel("Q2b_BL.xlsx") %>%
  janitor::clean_names()
```

```{r}
ggplot(baseline) + 
  geom_histogram(aes(x = age))
```

```{r}
summary(baseline)
```

*Simple randomization*

- Use logistic regression to find odds ratio

- 1 - exp(B) = Vaccine efficacy

- Want to know if Beta >= log(1 - 0.3). Use Wald's test with W = Beta - log(0.7)/se(Beta) and critical value z_0.05



# Question b

```{r}
baseline =  read_excel("Q2b_BL.xlsx", col_types = "numeric")

study =  read_excel("Q2b.xlsx", col_types = "numeric")

df_2 = left_join(baseline, study) %>%
  mutate(SEX = factor(SEX),
         GROUP = factor(GROUP),
         TIME = factor(TIME))
```


```{r}
naniar::mcar_test(df_2)
```


```{r}

# Fit logistic mixed effects model with random intercepts and random slopes
model <- glmer(SAE ~ TIME + AGE + SEX + GROUP +
                 (1 | ID/SITE) + GROUP*TIME, data = df_2, family = binomial)

# View model summary
summary(model)

#Use ANOVA to check if interaction term is necessary
```


```{r}

# Fit logistic mixed effects model with random intercepts and random slopes
modelt1 <- df_2 %>%
  filter(TIME == 1) %>%
  glm(SAE ~ AGE + SEX + GROUP, data = ., family = binomial)

# View model summary
summary(modelt1)

# Fit logistic mixed effects model with random intercepts and random slopes
modelt2 <- df_2 %>%
  filter(TIME == 2) %>%
  glm(SAE ~ AGE + SEX + GROUP, data = ., family = binomial)

# View model summary
summary(modelt2)


# Fit logistic mixed effects model with random intercepts and random slopes
modelt3 <- df_2 %>%
  filter(TIME == 3) %>%
  glm(SAE ~ AGE + SEX + GROUP, data = ., family = binomial)

# View model summary
summary(modelt3)

```


# Question c

```{r}
df_3 = read_excel("Q2c.xlsx", col_types = "numeric")
```

```{r}
table(df_3$InfectionTime, df_3$Infection)
```


```{r}
ggplot(data = df_3) + 
  geom_boxplot(aes(x = LastFUTime, y = factor(Infection)))
```

Checking if people got infected before second shot:

```{r}
df_3 %>%
  filter(EnrollmentTime < LastFUTime)
```

My understanding:

Patients were initially in first study (they can't have had COVID before then). They were followed up for some time.

Then, They get enrolled

They got the shot at some point after the follow up and enrollment.

They may have gotten infected after the vaccine (or otherwise missed to censoring)


```{r}
survfit(Surv(InfectionTime, event = Infection) ~ 1, conf.type='log', data = df_3)
```

```{r}
df_3 = df_3 %>%
  mutate(final_time = if_else(Infection == 0, LastFUTime, InfectionTime))

survfit(Surv(EnrollmentTime, final_time, Infection) ~ 1, 
        conf.type='log', data = df_3) %>%
  ggsurvfit()

survfit(Surv(EnrollmentTime, final_time, Infection) ~ 1, 
        conf.type='log', data = df_3) %>% summary(times = 365)
```

Note it does not reach 0.5

Using Cox model

```{r}
model = coxph(Surv(EnrollmentTime, final_time, Infection) ~ 1, 
         data = df_3, ties = "breslow")
```

Probability of survival after 12 months: Those that survived 12 months / those at risk (everybody in the dataset is at risk)

```{r}
surv_fit = survfit(Surv(EnrollmentTime, final_time, Infection) ~ 1, 
         data = df_3) 
  
1 - summary(surv_fit, times = 365)$surv

# Do "1 -" here
```
Mean:

```{r}
summary(surv_fit)$table["mean"]
```

```{r}
library(survRM2)

df_3$group <- 1  # Everyone in the same group

# Run rmst2 with the dummy group
rmst_result <- suppressMessages(
  suppressWarnings(
  rmst2(time = df_3$final_time,
                     status = df_3$Infection,
                     arm = df_3$group,
                     tau = 365)))  # Truncation time, e.g., 12 months

# Extract RMST estimate
rmst_result$RMST.arm1$est
```

Median with KM:

```{r}
summary(surv_fit)$table["median"]

surv_fit$median
```


```{r}
library(flexsurv)
# Fit a Weibull model with left truncation
weib_flex <- flexsurvreg(Surv(EnrollmentTime, final_time, Infection) ~ 1,
                         data = df_3,
                         dist = "weibull")

# Median survival time from the model
summary(weib_flex, type = "median")

summary(weib_flex, type = "mean")

exp_model <- flexsurvreg(Surv(EnrollmentTime, final_time, Infection) ~ 1,
                         data = df_3,
                         dist = "exponential")

# View model summary
summary(exp_model, type = "median")

summary(exp_model, type = "mean")
```
