---
title: "Project 4"
author: "Yuying Lu"
date: "2025-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# Main Objectives

-   

    a)  Comprehensive statistical analysis

    -   

        1)  Randomization procedure

    -   

        2)  Analytic approach for primary analysis

    -   

        3)  Sample size calculation

-   

    b)  Whether the vaccine group has greater odds of having SAE at any of the three assessment time points

-   

    c)  Other Interests

    -   

        1)  Probability of getting infected within 12 months with corresponding 95% confidence interval

    -   

        2)  Median time to infection, and the mean of the time to infection after the second shot




```{r}
library(tidyverse)
library(lme4)
library(lmerTest)
library(readxl)
library(ggplot2)
library(emmeans)
library(glmmTMB)
library(flexsurv)
library(sjPlot)
df_Q2bBL <- read_excel("Q2b_BL.xlsx") |> 
  janitor::clean_names()
df_Q2b <- read_excel("Q2b.xlsx")|> 
  janitor::clean_names()
df_Q2c <- read_excel("Q2c.xlsx", col_types = c("numeric", "numeric", "numeric","numeric", "numeric"))|> 
  janitor::clean_names()
```

## a)

### Randomized Clinical Trial

-   Recruit N participants who are seronegative at baseline.
    -   Stratified Sampling Method
    
-   Randomly split into placebo group and controlled groups with the same size (Double Blinded)
    -   Placebo group: get placebo (two injection)
    -   Controlled group: get vaccine (two injection)

-   Researchers follows the participants for 12 months and recorded their occurrence of symptomatic Covid-19 with onset at least 14 days


intent-to-treat Analysis

|     | D+  | D-  |Total|
|:---:|:---:|:---:|:---:|
| E+  | n1  | n2  | n1+n2|
| E-  | n3  | n4  | n3+n4| 


RR ={n1/(n1+n2)}/{n3/(n3+n4)}

H0: RR >= 0.7, H1: RR < 0.7

Test statistics: Wald Z = (log RR - log 0.7) / SE(log RR)

Reject H0 if Wald Z < -z_0.025 = -1.96
Report 95% CI for VE = 1-RR


### Assumption: 

Little loss of follow-up


### Power Analysis?

Provide 80% power to detect 60% vaccine effecacy without planned interim analysis


## b)

```{r}
df_b <- left_join(df_Q2bBL, df_Q2b, by = "id")
df_b
```

```{r}
tab <- df_b %>%
       group_by(time, group) %>%
       summarise(n = n(),
                 sae = sum(sae, na.rm = T),
                 risk = sae / n,
                 .groups = "drop")

print(tab)
```


### missingness


### Logistic Mixed Effect Model

Add site as random intercept (doesn't converge)

```{r}
SAE_mod <- df_b |> mutate(time = as.factor(time)) |> 
  glmer(sae ~ sex + age + time + group + (1 | id) + (1 | site),
            family = binomial(link = "logit"), 
            data   = _)
summary(SAE_mod)        
sjPlot::tab_model(SAE_mod, transform = "exp")  
```

Consider interaction between time and group

```{r}
SAE_mod_int <- df_b |> mutate(time = as.factor(time)) |> 
  glmer(sae ~ sex + age + time*group + (1 | id/site) ,
            family = binomial(link = "logit"), 
            data   = _)

summary(SAE_mod_int)
sjPlot::tab_model(SAE_mod_int, transform = "exp") 


```

### GEE ? Sensitivity Analysis

```{r}
library(geepack)
gee <- geeglm(sae ~ group * time + sex + age,
              id      = id,
              family  = binomial,
              corstr  = "exchangeable",
              data    = df_b)

summary(gee)
```


```{r}
library(emmeans)
emm <- emmeans(gee, ~ group | time, type = "response")
pairs(emm, by = "time")  # gives OR and 95% CI
```

## c)


```{r}
df_c = left_join(df_Q2c, df_Q2bBL, by = "id")
df_c
```



```{r}
mean(df_c$infection_time, na.rm = T)

mean(is.na(df_c$infection_time))
```

### KM

```{r,fig.width=6, fig.height=4}
library(survival)
library(survminer)
library(patchwork)
#library(pec)
#library(rms)
df_surv =  df_c |> 
       mutate(time = ifelse(infection == 1, infection_time,last_fu_time),
              status = infection)

km  <- survfit(Surv(enrollment_time, time, status) ~ 1, data = df_surv)
plot(km)
ggsurvplot(km, conf.int = T, data = df_surv, pval = TRUE, ggtheme = theme_minimal(),
           title = "KM Survival Curves")
# S(365) is KM survival prob at 12 months; 1‑S is infection probability

prob12 <- 1 - summary(km, times = 365)$surv
ci12   <- c(1 - summary(km, times = 365)$upper, 1 - summary(km, times = 365)$lower)

cat(sprintf("Probability of infection within 12 months = %.2f%% (95%% CI %.2f – %.2f%%)\n",
            100*prob12, 100*ci12[2], 100*ci12[1]))
```



```{r,fig.width=6, fig.height=4}
g <- ggsurvplot(
        km, conf.int = TRUE, pval = TRUE, data = df_surv,
        ggtheme = theme_minimal(), title = "KM Survival Curves"
     )

## 3. Grab the KM estimate and 95% CI at 365 days
s365      <- summary(km, times = 365)$surv
ci_lower  <- summary(km, times = 365)$lower
ci_upper  <- summary(km, times = 365)$upper

## 4. Add the vertical line, point, error bar, and label
g$plot <- g$plot +
  geom_vline(xintercept = 365, linetype = "dashed", colour = "red") +
  geom_point(aes(x = 365, y = s365), colour = "black", size = 3) +
  geom_errorbar(
      aes(x = 365, ymin = ci_lower, ymax = ci_upper),
      width = 0, colour = "black", linewidth = 0.8
  ) +
  annotate(
      "text", x = 367, y = s365,
      label = sprintf("  S(365) = %.2f\n95%% CI: %.2f–%.2f",
                      s365, ci_lower, ci_upper),
      hjust = -0.05, vjust = 0.15, colour = "black"
  )

## 5. Display
print(g)

est_km = summary(km)$table['rmean']
se_km = summary(km)$table['se(rmean)']
```




```{r}
fit_flex <- flexsurvreg(Surv(enrollment_time, time, status) ~ 1,
                        data = df_surv, dist = "exp")  # same as exponential
print(fit_flex, ci = TRUE)
```

Check the exponential assumption:

```{r}
ggsurv <- survfit(Surv(enrollment_time, time, status) ~ 1,
                        data = df_surv)
plot(log(ggsurv$time), log(ggsurv$surv), type = "s")
```
```{r}

weib_flex <- flexsurvreg(Surv(enrollment_time, time, status) ~ 1,
                         data = df_surv,
                         dist = "weibull")

shape_k <- exp(coefficients(weib_flex)[1])
scale_lambda <- exp(coefficients(weib_flex)[2])
scale_lambda * (log(2))^(1/shape_k)

# Median survival time from the model

summary(weib_flex, type = "median")
summary(weib_flex, type = "mean")

exp_model <- flexsurvreg(Surv(enrollment_time, time, status) ~ 1,
                         data = df_surv,
                         dist = "exponential")

# View model summary
lambda = exp(coefficients(exp_model))
summary(exp_model, type = "median") 
summary(exp_model, type = "mean")

data_1 = 
  data.frame(Distribution = c(rep("Exp",2),rep("Weibull",2),"KM"),
           Time_type = c("median","mean","median","mean","mean"),
           dplyr::bind_rows(summary(exp_model, type = "median"),
      summary(exp_model, type = "mean"),
      summary(weib_flex, type = "median"),
      summary(weib_flex, type = "mean"),
      data.frame(est = est_km, lcl = est_km-1.96*se_km,ucl = est_km+1.96*se_km)))


data_1


```

```{r}
pred_exp <- summary(exp_model,t = 356, type = "survival")
pred_weib <- summary(weib_flex,t = 356, type = "survival")
pred_log <- summary(log_lgt,t = 356, type = "survival")
pred_gamma <- summary(gamma_mod,t = 356, type = "survival")
s365      <- 1- summary(km, times = 365)$surv
ci_lower  <- 1- summary(km, times = 365)$lower
ci_upper  <- 1- summary(km, times = 365)$upper

1-pred[[1]][c("est", "ucl", "lcl")]

data.frame(Distribution = c("Exp","Weibull","Log-logistic","Gamma","KM"),
           dplyr::bind_rows(1-pred_exp[[1]][c("est", "ucl", "lcl")],
      1-pred_weib[[1]][c("est", "ucl", "lcl")],
      1-pred_log[[1]][c("est", "ucl", "lcl")],
      1-pred_gamma[[1]][c("est", "ucl", "lcl")],
      data_frame(est=s365, ucl = ci_upper, lcl=ci_lower)))




```


```{r}
gamma_mod <- flexsurvreg(Surv(enrollment_time, time, status) ~ 1,
                         data = df_surv,
                         dist = "gamma")


rate_gamma          <- gamma_mod$res[,"est"]["rate"]         # λ
shape_k_gamma       <- gamma_mod$res[,"est"]["shape"]  

mean_gamma = shape_k_gamma/rate_gamma

summary(gamma_mod, type = "median")

summary(gamma_mod, type = "mean")

log_lgt <- flexsurvreg(Surv(enrollment_time, time, status) ~ 1,
                         data = df_surv,
                         dist = "llogis")

kappa_log <- log_lgt$res["shape", "est"]
rate_log <- log_lgt$res["scale", "est"]
summary(log_lgt, type = "median") 
summary(log_lgt, type = "mean")

data_2 = 
  data.frame(Distribution = c(rep("Log-logistic",2),rep("Gamma",2)),
           Time_type = c("median","mean","median","mean"),
           dplyr::bind_rows(summary(log_lgt, type = "median"),
      summary(log_lgt, type = "mean"),
      summary(gamma_mod, type = "median"),
      summary(gamma_mod, type = "mean")))
```


```{r}
dplyr::bind_rows(data_1,data_2)
```


```{r}
km_df = summary(km, times = df_surv$time)
line_size = 0.7
survival_plot_A <- 
  ggplot(df_surv, aes(x = time)) +
  #stat_ecdf(geom = "step", aes(y = 1 - ..y.., color = "Empirical"), size = 1) +  
  stat_function(fun = function(x) exp(-lambda * x), 
                aes(color = "Exponential"), size = line_size )  +
  stat_function(fun = function(x) exp(-(x/scale_lambda)^shape_k),  
                aes(color = "Weibull"), size = line_size ) + 
  stat_function(fun = function(x) 1/(1+(x/rate_log)^kappa_log),  
                aes(color = "Log-logistic"), size = line_size ) + 
  stat_function(fun = function(x) pgamma(x, shape = shape_k_gamma, rate = rate_gamma, lower.tail = FALSE),  
                aes(color = "Gamma"), size = line_size ) + 
  geom_step(aes(y = km_df$surv, x = km_df$time, color = "Kaplan-Meier"), size = line_size )+
  scale_color_manual(name = "Method", 
                     limits = c("Kaplan-Meier", "Exponential", "Weibull", "Log-logistic", "Gamma"),
                     values = c("Exponential" = "#D16103", 
                                "Weibull" = "#22A54C",
                                "Log-logistic" = "#4E84C4",
                                "Gamma" = "#BC0907",
                                "Kaplan-Meier" = "#BC89A7")) +  
  labs(title = "Survival Function Comparison",
       x = "Time",
       y = "Survival Probability")  +
  theme_minimal()


survival_plot_A
```



# fit other distribution


```{r}
cox_mod<- coxph(Surv(enrollment_time, time, status) ~ sex + age , data = df_surv, x= TRUE)
# Print summary of the model
summary(cox_mod)
cox.zph(cox_mod)
```



```{r}
baseline_hazard <- basehaz(cox_mod, centered = FALSE)
head(baseline_hazard)

plot(baseline_hazard$time, baseline_hazard$hazard, type = "s",
     xlab = "Time", ylab = "Cumulative Baseline Hazard",
     main = "Baseline Hazard Function", col = "#4E84C4", lwd = 2)


# Plot using ggplot2
cum_plot_B = ggplot(baseline_hazard, aes(x = time, y = hazard)) +
  geom_step(color = "#4E84C4", size = 1.2) +  # Step plot for cumulative hazard
  labs(title = expression("Baseline Hazard Function " ~ H[0](t)),
       x = "Time",
       y = "Cumulative Baseline Hazard") +
  theme_minimal()

baseline_hazard$survival <- exp(-baseline_hazard$hazard)
baseline_hazard$time
# Plot the survival function using ggplot2
surv_plot_B = ggplot(baseline_hazard, aes(x = time, y = survival)) +
  geom_step(color = "#4E84C4", size = 1.2) +  # Step plot for survival function
  labs(title = expression("Baseline Survival Function " ~ S[0](t)),
       x = "Time",
       y = "Survival Probability") +
  theme_minimal()
surv_plot_B
```
